Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows

### 2024-10-21 manuscript/manuscript.Rmd
update ms.

The hazard models in this tutorial assume that there is one event of interest. For RT data, this event constitutes a single transition between an "idle" state and a "responded" state. However, in certain situations, more than one event of interest might exist. For example, in a medical or health-related context, an individual might transition back and forth between a "healthy" state and a "depressed" state, before a final "death" state. When you have data on the timing of these transitions, one can apply multi-state models, which generalize survival analysis to transitions between three or more states [@SteeleF2004].
Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial. Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial [@@allisonSurvivalAnalysisUsing2010]. For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence. Note that the effect of a time-varying predictor (e.g., an occipital EEG signal) can itself vary over time.

>>>>

The hazard models in this tutorial assume that there is one event of interest. For RT data, this event constitutes a single transition between an "idle" state and a "responded" state. However, in certain situations, more than one event of interest might exist. For example, in a medical or health-related context, an individual might transition back and forth between a "healthy" state and a "depressed" state, before a final "death" state. When you have data on the timing of these transitions, one can apply multi-state models, which generalize survival analysis to transitions between three or more states [@SteeleF2004].
Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial. Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial [@allisonSurvivalAnalysisUsing2010]. For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence. Note that the effect of a time-varying predictor (e.g., an occipital EEG signal) can itself vary over time.

### 2024-10-21 manuscript/manuscript.tex
update ms.

The hazard models in this tutorial assume that there is one event of interest. For RT data, this event constitutes a single transition between an ``idle'' state and a ``responded'' state. However, in certain situations, more than one event of interest might exist. For example, in a medical or health-related context, an individual might transition back and forth between a ``healthy'' state and a ``depressed'' state, before a final ``death'' state. When you have data on the timing of these transitions, one can apply multi-state models, which generalize survival analysis to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial. Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial (@ Allison, 2010). For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence. Note that the effect of a time-varying predictor (e.g., an occipital EEG signal) can itself vary over time.

>>>>

The hazard models in this tutorial assume that there is one event of interest. For RT data, this event constitutes a single transition between an ``idle'' state and a ``responded'' state. However, in certain situations, more than one event of interest might exist. For example, in a medical or health-related context, an individual might transition back and forth between a ``healthy'' state and a ``depressed'' state, before a final ``death'' state. When you have data on the timing of these transitions, one can apply multi-state models, which generalize survival analysis to transitions between three or more states (Steele, Goldstein, \& Browne, 2004).
Also, the predictor variables in this tutorial are time-invariant, i.e., their value did not change over the course of a trial. Thus, another extension is to include time-varying predictors, i.e., predictors whose value can change across the time bins within a trial (Allison, 2010). For example, when gaze position is tracked during a visual search trial, the gaze-target distance will vary during a trial when the eyes move around before a manual response is given; shorter gaze-target distances should be associated with a higher hazard of response occurrence. Note that the effect of a time-varying predictor (e.g., an occipital EEG signal) can itself vary over time.
